\markboth{\bibname}{}
\phantomsection  % so hyperref creates bookmarks
\addcontentsline{toc}{chapter}{\bibname}

\begin{thebibliography}{99}

% from pre-face
% ====================

\bibitem{fp-pearls}
Richard Bird. ``Pearls of functional algorithm design''. Cambridge University Press; 1 edition (November 1, 2010). ISBN-10: 0521513383. pp1 - pp6.

\bibitem{Bentley}
Jon Bentley. ``Programming Pearls(2nd Edition)''. Addison-Wesley Professional; 2 edition (October 7, 1999). ISBN-13: 978-0201657883 （中文版：《编程珠玑》）

\bibitem{okasaki-book}
Chris Okasaki. ``Purely Functional Data Structures''. Cambridge university press, (July 1, 1999), ISBN-13: 978-0521663502

\bibitem{CLRS}
Thomas H. Cormen, Charles E. Leiserson, Ronald L. Rivest and Clifford Stein. ``Introduction to Algorithms, Second Edition''. The MIT Press, 2001. ISBN: 0262032937. （中文版：《算法导论》）

% from BST
% ==================

\bibitem{okasaki-blog}
Chris Okasaki. ``Ten Years of Purely Functional Data Structures''. \url{http://okasaki.blogspot.com/2008/02/ten-years-of-purely-functional-data.html}

\bibitem{sgi-stl}
SGI. ``Standard Template Library Programmer's Guide''. \url{http://www.sgi.com/tech/stl/}

% not available any more
%\bibitem{literal-program}
%\url{http://en.literateprograms.org/Category:Binary_search_tree}

\bibitem{wiki-fold}
Wikipedia. ``Fold(high-order function)''. \url{https://en.wikipedia.org/wiki/Fold_(higher-order_function)}

\bibitem{func-composition}
Wikipedia. ``Function Composition''. \url{https://en.wikipedia.org/wiki/Function_composition}

\bibitem{curry}
Wikipedia. ``Partial application''. \url{https://en.wikipedia.org/wiki/Partial_application}

\bibitem{learn-haskell}
Miran Lipovaca. ``Learn You a Haskell for Great Good! A Beginner's Guide''. No Starch Press; 1 edition April 2011, 400 pp. ISBN: 978-1-59327-283-8

% from isort
% ============================
\bibitem{wiki-bubble-sort}
Wikipedia. ``Bubble sort''. \url{https://en.wikipedia.org/wiki/Bubble_sort}

\bibitem{Knuth-V3}
Donald E. Knuth. ``The Art of Computer Programming, Volume 3: Sorting and Searching (2nd Edition)''. Addison-Wesley Professional; 2 edition (May 4, 1998) ISBN-10: 0201896850 ISBN-13: 978-0201896855

% from rbtree
% ===========================
\bibitem{okasaki}
Chris Okasaki. ``FUNCTIONAL PEARLS Red-Black Trees in a Functional Setting''. J. Functional Programming. 1998

\bibitem{wiki-rbt}
Wikipedia. ``Red-black tree''. \url{https://en.wikipedia.org/wiki/Red-black_tree}

\bibitem{lyn}
Lyn Turbak. ``Red-Black Trees''. \url{http://cs.wellesley.edu/~cs231/fall01/red-black.pdf} Nov. 2, 2001.

\bibitem{rosetta}
Rosetta Code. ``Pattern matching''. \url{http://rosettacode.org/wiki/Pattern_matching}

% from avl tree
% =====================
\bibitem{hackage-avl}
Hackage. ``Data.Tree.AVL''. \url{http://hackage.haskell.org/packages/archive/AvlTree/4.2/doc/html/Data-Tree-AVL.html}

\bibitem{wiki-avl}
Wikipedia. ``AVL tree''. \url{https://en.wikipedia.org/wiki/AVL_tree}

\bibitem{TFATP}
Guy Cousinear, Michel Mauny. ``The Functional Approach to Programming''. Cambridge University Press; English Ed edition (October 29, 1998). ISBN-13: 978-0521576819

\bibitem{py-avl}
Pavel Grafov. ``Implementation of an AVL tree in Python''. \url{http://github.com/pgrafov/python-avl-tree}

% from radix tree (trie and Patricia)
% =======================================

\bibitem{okasaki-int-map}
Chris Okasaki and Andrew Gill. ``Fast Mergeable Integer Maps''. Workshop on ML, September 1998, pages 77-86.

\bibitem{patricia-morrison}
D.R. Morrison, ``PATRICIA -- Practical Algorithm To Retrieve  Information Coded In Alphanumeric", Journal of the ACM, 15(4), October 1968, pages 514-534.

\bibitem{wiki-suffix-tree}
Wikipedia. ``Suffix Tree''. \url{https://en.wikipedia.org/wiki/Suffix_tree}

\bibitem{wiki-trie}
Wikipedia. ``Trie''. \url{https://en.wikipedia.org/wiki/Trie}

\bibitem{wiki-t9}
Wikipedia. ``T9 (predictive text)''. \url{https://en.wikipedia.org/wiki/T9_(predictive_text)}

\bibitem{wiki-predictive-text}
Wikipedia. ``Predictive text''. \url{https://en.wikipedia.org/wiki/Predictive_text}

% from suffix tree
% ===================================

\bibitem{ukkonen95}
Esko Ukkonen. ``On-line construction of suffix trees''. Algorithmica 14 (3): 249--260. doi:10.1007/BF01206331. \url{http://www.cs.helsinki.fi/u/ukkonen/SuffixT1withFigs.pdf}

\bibitem{weiner73}
Weiner, P. ``Linear pattern matching algorithms'', 14th Annual IEEE Symposium on Switching and Automata Theory, pp. 1-11, doi:10.1109/SWAT.1973.13

\bibitem{ukkonen-presentation}
Esko Ukkonen. ``Suffix tree and suffix array techniques for pattern analysis in strings''. \url{http://www.cs.helsinki.fi/u/ukkonen/Erice2005.ppt}

\bibitem{trivial-stree-java}
Suffix Tree (Java). \url{http://en.literateprograms.org/Suffix_tree_(Java)}

\bibitem{GieKur97}
Robert Giegerich and Stefan Kurtz. ``From Ukkonen to McCreight and Weiner: A Unifying View of Linear-Time Suffix Tree Construction''. Science of Computer Programming 25(2-3):187-218, 1995. \url{http://citeseer.ist.psu.edu/giegerich95comparison.html}

\bibitem{GieKur95}
Robert Giegerich and Stefan Kurtz. ``A Comparison of Imperative and Purely Functional Suffix Tree Constructions''. Algorithmica 19 (3): 331--353. doi:10.1007/PL00009177. \url{http://www.zbh.uni-hamburg.de/pubs/pdf/GieKur1997.pdf}

\bibitem{Hackage-STree}
Bryan O'Sullivan. ``suffixtree: Efficient, lazy suffix tree implementation''. \url{http://hackage.haskell.org/package/suffixtree}

\bibitem{plt-stree}
Danny. \url{http://hkn.eecs.berkeley.edu/~dyoo/plt/suffixtree/}

\bibitem{Gusfield-book}
Dan Gusfield. ``Algorithms on Strings, Trees and Sequences Computer Science and Computational Biology''. Cambridge University Press; 1 edition (May 28, 1997) ISBN: 9780521585194

\bibitem{lallison-stree}
Lloyd Allison. ``Suffix Trees''. \url{http://www.allisons.org/ll/AlgDS/Tree/Suffix/}

\bibitem{ukkonen-lec}
Esko Ukkonen. ``Suffix tree and suffix array techniques for pattern analysis in strings''. \url{http://www.cs.helsinki.fi/u/ukkonen/Erice2005.ppt}

\bibitem{ukkonen-search}
Esko Ukkonen ``Approximate string-matching over suffix trees''. Proc. CPM 93. Lecture Notes in Computer Science 684, pp. 228-242, Springer 1993. \url{http://www.cs.helsinki.fi/u/ukkonen/cpm931.ps}

% from B-tree
% =====================

\bibitem{wiki-b-tree}
Wikipeida. ``B-tree''. \url{https://en.wikipedia.org/wiki/B-tree}

% from binary healp
% =====================

\bibitem{wiki-heap}
Wikipedia. ``Heap (data structure)''. \url{https://en.wikipedia.org/wiki/Heap_(data_structure)}

\bibitem{wiki-heapsort}
Wikipedia. ``Heapsort''. \url{https://en.wikipedia.org/wiki/Heapsort}

\bibitem{rosetta-heapsort}
Rosetta Code. ``Sorting algorithms/Heapsort''.  \url{http://rosettacode.org/wiki/Sorting_algorithms/Heapsort}

\bibitem{wiki-leftist-tree}
Wikipedia. ``Leftist Tree''. \url{https://en.wikipedia.org/wiki/Leftist_tree}

\bibitem{brono-book}
Bruno R. Preiss. Data Structures and Algorithms with Object-Oriented Design Patterns in Java. \url{http://www.brpreiss.com/books/opus5/index.html}

\bibitem{TAOCP-bheap}
Donald E. Knuth. ``The Art of Computer Programming. Volume 3: Sorting and Searching.''. Addison-Wesley Professional;
2nd Edition (October 15, 1998). ISBN-13: 978-0201485417. Section 5.2.3 and 6.2.3

\bibitem{wiki-skew-heap}
Wikipedia. ``Skew heap''. \url{https://en.wikipedia.org/wiki/Skew_heap}

\bibitem{self-adjusting-heaps}
Sleator, Daniel Dominic; Jarjan, Robert Endre. ``Self-adjusting heaps'' SIAM Journal on Computing 15(1):52-69. doi:10.1137/0215004 ISSN 00975397 (1986)

\bibitem{wiki-splay-tree}
Wikipedia. ``Splay tree''. \url{https://en.wikipedia.org/wiki/Splay_tree}

\bibitem{self-adjusting-trees}
Sleator, Daniel D.; Tarjan, Robert E. (1985), ``Self-Adjusting Binary Search Trees'', Journal of the ACM 32(3):652 - 686, doi: 10.1145/3828.3835

\bibitem{NIST}
NIST, ``binary heap''. \url{http://xw2k.nist.gov/dads//HTML/binaryheap.html}

% from selection sort
% ====================
\bibitem{TAOCP}
Donald E. Knuth. ``The Art of Computer Programming, Volume 3: Sorting and Searching (2nd Edition)''. Addison-Wesley Professional; 2 edition (May 4, 1998) ISBN-10: 0201896850 ISBN-13: 978-0201896855

\bibitem{wiki-sweak-order}
Wikipedia. ``Strict weak order''. \url{https://en.wikipedia.org/wiki/Strict_weak_order}

\bibitem{wiki-wc}
Wikipedia. ``FIFA world cup''. \url{https://en.wikipedia.org/wiki/FIFA_World_Cup}

% from k-help
% ====================
\bibitem{K-ary-tree}
Wikipedia. ``K-ary tree''. \url{https://en.wikipedia.org/wiki/K-ary_tree}

\bibitem{wiki-pascal-triangle}
Wikipedia, ``Pascal's triangle''. \url{https://en.wikipedia.org/wiki/Pascal's_triangle}

\bibitem{hackage-fibq}
Hackage. ``An alternate implementation of a priority queue based on a Fibonacci heap.'', \url{http://hackage.haskell.org/packages/archive/pqueue-mtl/1.0.7/doc/html/src/Data-Queue-FibQueue.html}

\bibitem{okasaki-fibh}
Chris Okasaki. ``Fibonacci Heaps.'' \url{http://darcs.haskell.org/nofib/gc/fibheaps/orig}

\bibitem{pairing-heap}
Michael L. Fredman, Robert Sedgewick, Daniel D. Sleator, and Robert E. Tarjan. ``The Pairing Heap: A New Form of Self-Adjusting Heap'' Algorithmica (1986) 1: 111-129.

% from queue
% =====================
\bibitem{PODC96}
Maged M. Michael and Michael L. Scott. ``Simple, Fast, and Practical Non-Blocking and Blocking Concurrent Queue Algorithms''. \url{http://www.cs.rochester.edu/research/synchronization/pseudocode/queues.html}

\bibitem{SutterDDJ}
Herb Sutter. ``Writing a Generalized Concurrent Queue''. Dr. Dobb's Oct 29, 2008. \url{http://drdobbs.com/cpp/211601363?pgno=1}

\bibitem{wiki-tail-call}
Wikipedia. ``Tail-call''. \url{https://en.wikipedia.org/wiki/Tail_call}

\bibitem{recursion}
Wikipedia. ``Recursion (computer science)''. \url{https://en.wikipedia.org/wiki/Recursion_(computer_science)#Tail-recursive_functions}

\bibitem{SICP}
Harold Abelson, Gerald Jay Sussman, Julie Sussman. ``Structure and Interpretation of Computer Programs, 2nd Edition''. MIT Press, 1996, ISBN 0-262-51087-1 (中文版：裘宗燕 译《计算机程序的构造和解释》)

% sequence
% =====================

\bibitem{okasaki-ralist}
Chris Okasaki. ``Purely Functional Random-Access Lists''. Functional Programming Languages and Computer Architecture, June 1995, pages 86-95.

\bibitem{finger-tree-2006}
Ralf Hinze and Ross Paterson. ``Finger Trees: A Simple General-purpose Data Structure,'' in Journal of Functional Programming 16:2 (2006), pages 197-217. \url{http://www.soi.city.ac.uk/~ross/papers/FingerTree.html}

\bibitem{finger-tree-1977}
Guibas, L. J., McCreight, E. M., Plass, M. F., Roberts, J. R. (1977), "A new representation for linear lists". Conference Record of the Ninth Annual ACM Symposium on Theory of Computing, pp. 49-60.

\bibitem{hackage-ftr}
Generic finger-tree structure. \url{http://hackage.haskell.org/packages/archive/fingertree/0.0/doc/html/Data-FingerTree.html}

\bibitem{mtf-wiki}
Wikipedia. ``Move-to-front transform''. \url{https://en.wikipedia.org/wiki/Move-to-front_transform}

% dc sort
% ===================

\bibitem{qsort-impl}
Robert Sedgewick. ``Implementing quick sort programs''. Communication of ACM. Volume 21, Number 10. 1978. pp.847 - 857.

\bibitem{3-way-part}
Jon Bentley, Douglas McIlroy. ``Engineering a sort function''. Software Practice and experience VOL. 23(11), 1249-1265 1993.

\bibitem{opt-qs}
Robert Sedgewick, Jon Bentley. ``Quicksort is optimal''. \url{http://www.cs.princeton.edu/~rs/talks/QuicksortIsOptimal.pdf}

\bibitem{algo-fp}
Fethi Rabhi, Guy Lapalme. ``Algorithms: a functional programming approach''. Second edition. Addison-Wesley, 1999. ISBN: 0201-59604-0

\bibitem{slpj-book-1987}
Simon Peyton Jones. ``The Implementation of functional programming languages''. Prentice-Hall International, 1987. ISBN: 0-13-453333-X

\bibitem{msort-in-place}
Jyrki Katajainen, Tomi Pasanen, Jukka Teuhola. ``Practical in-place mergesort''. Nordic Journal of Computing, 1996.

\bibitem{sort-deriving}
Jos\`{e} Bacelar Almeida and Jorge Sousa Pinto. ``Deriving Sorting Algorithms''. Technical report, Data structures and Algorithms. 2008.

\bibitem{para-msort}
Cole, Richard (August 1988). ``Parallel merge sort''. SIAM J. Comput. 17 (4): 770-785. doi:10.1137/0217049. (August 1988)

\bibitem{para-qsort}
Powers, David M. W. ``Parallelized Quicksort and Radixsort with Optimal Speedup'', Proceedings of International Conference on Parallel Computing Technologies. Novosibirsk. 1991.

\bibitem{wiki-qs}
Wikipedia. ``Quicksort''. \url{https://en.wikipedia.org/wiki/Quicksort}

\bibitem{wiki-total-order}
Wikipedia. ``Total order''. \url{http://en.wokipedia.org/wiki/Total_order}

\bibitem{wiki-harmonic}
Wikipedia. ``Harmonic series (mathematics)''. \url{https://en.wikipedia.org/wiki/Harmonic_series_(mathematics)}

% from search
% =====================

\bibitem{median-of-median}
M. Blum, R.W. Floyd, V. Pratt, R. Rivest and R. Tarjan, "Time bounds for selection," J. Comput. System Sci. 7 (1973) 448-461.

\bibitem{saddle-back}
Edsger W. Dijkstra. ``The saddleback search''. EWD-934. 1985. \url{https://www.cs.utexas.edu/users/EWD/ewd09xx/EWD934.PDF}.

\bibitem{boyer-moore-majority}
Robert Boyer, and Strother Moore. ``MJRTY - A Fast Majority Vote Algorithm''. Automated Reasoning: Essays in Honor of Woody Bledsoe, Automated Reasoning Series, Kluwer Academic Publishers, Dordrecht, The Netherlands, 1991, pp. 105-117.

\bibitem{count-min-sketch}
Cormode, Graham; S. Muthukrishnan (2004). ``An Improved Data Stream Summary: The Count-Min Sketch and its Applications''. J. Algorithms 55: 29-38.

\bibitem{kmp}
Knuth Donald, Morris James H., jr, Pratt Vaughan. ``Fast pattern matching in strings''. SIAM Journal on Computing 6 (2): 323-350. 1977.

\bibitem{boyer-moore}
Robert Boyer, Strother Moore. ``A Fast String Searching Algorithm''. Comm. ACM (New York, NY, USA: Association for Computing Machinery) 20 (10): 762-772. 1977

\bibitem{boyer-moore-horspool}
R. N. Horspool. ``Practical fast searching in strings''. Software - Practice \& Experience 10 (6): 501-506. 1980.

\bibitem{wiki-boyer-moore}
Wikipedia. ``Boyer-Moore string search algorithm''. \url{https://en.wikipedia.org/wiki/Boyer-Moore_string_search_algorithm}

\bibitem{wiki-8-queens}
Wikipedia. ``Eight queens puzzle''. \url{https://en.wikipedia.org/wiki/Eight_queens_puzzle}

\bibitem{how-to-solve-it}
George P\'{o}lya. ``How to solve it: A new aspect of mathematical method''. Princeton University Press(April 25, 2004). ISBN-13: 978-0691119663

\bibitem{Huffman}
Wikipedia. ``David A. Huffman''. \url{https://en.wikipedia.org/wiki/David_A._Huffman}

% appendix list
% =====================

\bibitem{moderncxx}
Andrei Alexandrescu. ``Modern C++ design: Generic Programming and Design Patterns Applied''. Addison Wesley February 01, 2001, ISBN 0-201-70431-5

\bibitem{mittype}
Benjamin C. Pierce. ``Types and Programming Languages''. The MIT Press, 2002. ISBN:0262162091

\bibitem{erlang}
Joe Armstrong. ``Programming Erlang: Software for a Concurrent World''. Pragmatic Bookshelf; 1 edition (July 18, 2007). ISBN-13: 978-1934356005

\bibitem{sgi-stl-transform}
SGI. ``transform''. \url{http://www.sgi.com/tech/stl/transform.html}

\bibitem{poj-drunk-jailer}
ACM/ICPC. ``The drunk jailer.'' Peking University judge online for ACM/ICPC. \url{http://poj.org/problem?id=1218}.

\bibitem{Haskell-wiki}
Haskell wiki. ``Haskell programming tips''. 4.4 Choose the appropriate fold. \url{http://www.haskell.org/haskellwiki/Haskell_programming_tips}

\bibitem{wiki-dot-product}
Wikipedia. ``Dot product''. \url{https://en.wikipedia.org/wiki/Dot_product}

\bibitem{unplugged}
Xinyu LIU. ``Isomorphism - mathematics of programming''. \url{https://github.com/liuxinyu95/unplugged}

\end{thebibliography}
